A longitudinal dataset of childhood gut microbiome linked with cognitive evaluation

### 2021-05-07 manuscript/main.tex
finalize

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
% \usepackage{hyperref}
% \usepackage{academicons}
% \usepackage{xcolor}

@@

General participant demographics are provided in Tables \ref{tab:agestats}, \ref{tab:race}, and \ref{tab:breastfeeding} and Figure \ref{fig:data}.
Complete metadata are available in Data Records (see below), with children being representative of the RI population.
As a broad background, children in the RESONANCE cohort were born full-term (>37 weeks gestation)
with height and weight normal for gestational age, and from uncomplicated singleton pregnancies.
Children with known major risk factors for developmental abnormalities at enrollment were excluded.
In addition to screening at the time of enrollment,
on-going screening for worrisome behaviors using validated tools was performed
to identify at-risk children and remove them from subsequent analysis.

@@

Taxonomic profiles obtained from metagenomic sequencing using MetaPhlAn are available on OSF.io (\verb|metaphlan_profiles.tar.gz|).
A single tab-separated values (tsv) file for each sample is provided,
with columns corresponding to the taxon name, NCBI Taxonomic ID, relative abundance, and additional species.
These tables may be merged using the \verb|metaphlan_merge_tables| command line utility provided by the MetaPhlAn package,
or using other software that can read and join tables.
A convenience function (\verb|taxonomic_profiles()|) for downloading and loading these tables in a julia session
is provided with \verb|ResonanceMicrobiome.jl|.
Functional profiles obtained from metagenomic sequencing using HUMAnN are available on OSF.io.
There are several different schemas for identifying protein-coding genes.
Initial mapping to UniRef90 \cite{UniProt_Consortium2010-ls} gene clusters was performed by HUMAnN,
with subsequent regrouping into Protein Family (pfam) \cite{Bateman2004-cz,El-Gebali2019-jh}, Kegg Orthology (KO) \cite{Kanehisa2000-rx,Kanehisa2008-gg},
and level 4 Enzyme Commission (EC) \cite{Bairoch2000-ot} groups using the \verb|humann_regroup_tables| command-line utility.
A convenience function (\verb|functional_profiles()|) for downloading and loading these tables in a julia session
is provided with \verb|ResonanceMicrobiome.jl|.

@@

The corresponding author is responsible for providing a \href{https://www.nature.com/sdata/policies/editorial-and-publishing-policies#competing}{competing interests statement} on behalf of all authors of the paper.
This statement must be included in the submitted article file

@@

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
    \hline\hline
    \textbf{ } & \textbf{Mean} & \textbf{Median} & \textbf{Minimum} & \textbf{Maximum} & \textbf{Total} \\\hline
    1 and under & 0.45 & 0.32 & 0.08 & 1.0 & 131 \\
    1 to 2 & 1.52 & 1.55 & 1.01 & 1.99 & 53 \\
    2 and over & 6.59 & 6.27 & 2.01 & 15.29 & 309 \\\hline\hline
  \end{tabular}
\caption{
    Age statistics of children
}
\label{tab:agestats}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rr}
    \hline\hline
    \textbf{Race} & \textbf{Number} \\\hline
    Asian  & 9 \\
    Native American / Alaskan Native & 9 \\
    African American / Black & 17 \\
    Unknown & 44 \\
    Mixed & 93 \\
    Caucasian / White & 321 \\\hline\hline
\end{tabular}
\caption{
    Race / ethnicity of subjects
}
\label{tab:race}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rr}
    \hline\hline
    \textbf{Liquid diet} & \textbf{Number} \\\hline
    mixed & 14 \\
    exclusive formula & 35 \\
    exclusive breast & 48 \\
    missing & 392 \\\hline\hline
\end{tabular}
\caption{
    Breastfeeding counts
}
\label{tab:breastfeeding}
\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{../figures/05_data_summaries}
\caption{
    Graphical representations of subject and sample-associated metadata.
    % (a) Ages of children when samples were collected for all children
    % or just children under 2 years of age (b).
    % (c) Total number of samples from children or samples at each age (d)
    % that were exclussively breastfed,
    % exclussively formula fed, or with a mixed diet.
    % These values refer to the diet \emph{when on liquid diet},
    % not necessarily current diet.
    % (d) Race and ethnicity of participants.
    % (e) Maternal SES as measured by the Hollingshead Scale.
    % (f) Age-normalized BMI z-score of children when samples were collected.
    % (g) Age-normalized cognitive score of children when samples were collected.
    % (f) Subjects with longitudinal timepoints by first collection.
    % "Infant": birth to 1 year, "early childhood": 1 year to 6 years,
    % "middle childhood": 6 years to 12 years, "adolescent": older than 12 years.
}
\label{fig:data}
\end{figure}

>>>>

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lineno}
\usepackage{adjustbox}
\usepackage{multirow}
% \usepackage{hyperref}
% \usepackage{academicons}
% \usepackage{xcolor}

@@

General participant demographics are provided in Tables \ref{tab:demographics} \ref{tab:agestats} and Figure \ref{fig:data}.
Complete metadata are available in Data Records (see below), with children being representative of the RI population.
As a broad background, children in the RESONANCE cohort were born full-term (>37 weeks gestation)
with height and weight normal for gestational age, and from uncomplicated singleton pregnancies.
Children with known major risk factors for developmental abnormalities at enrollment were excluded.
In addition to screening at the time of enrollment,
on-going screening for worrisome behaviors using validated tools was performed
to identify at-risk children and remove them from subsequent analysis.

@@

Taxonomic profiles obtained from metagenomic sequencing using MetaPhlAn are available on OSF.io
as separate tab-separated values (tsv) files for each sample in \verb|metaphlan_profiles.tar.gz|.
A single table for each sample is provided,
with columns corresponding to the taxon name, NCBI Taxonomic ID, relative abundance, and additional species.
These tables may be merged using the \verb|metaphlan_merge_tables| command line utility provided by the MetaPhlAn package,
or using other software that can read and join tables.
A convenience function (\verb|taxonomic_profiles()|) for downloading and loading these tables in a julia session
is provided with \verb|ResonanceMicrobiome.jl|.
Functional profiles obtained from metagenomic sequencing using HUMAnN are available on OSF.io.
There are several different schemas for identifying protein-coding genes.
Initial mapping to UniRef90 \cite{UniProt_Consortium2010-ls} gene clusters was performed by HUMAnN,
with subsequent regrouping into Protein Family (pfam) \cite{Bateman2004-cz,El-Gebali2019-jh}, Kegg Orthology (KO) \cite{Kanehisa2000-rx,Kanehisa2008-gg},
and level 4 Enzyme Commission (EC) \cite{Bairoch2000-ot} groups using the \verb|humann_regroup_tables| command-line utility.
A convenience function (\verb|functional_profiles()|) for downloading and loading these tables in a julia session
is provided with \verb|ResonanceMicrobiome.jl|.

@@

The authors declare no competing interests.

@@

\begin{table}[ht]
\centering
\begin{tabular}{lrcc}
    \hline\hline
    \textbf{ } & \textbf{Category} & \textbf{Number} & \textbf{\%} \\\hline
    \multirow{6}{*}{Race} & Native American / Alaskan Native & 4 & 1.2\%\\
                          & Asian  & 6 & 1.7\%\\
                          & African American / Black & 14 & 4.1\%\\
                          & Unknown & 43 & 12.5\%\\
                          & Mixed & 64 & 18.6\%\\
                          & Caucasian / White & 213 & 61.9\%\\\hline
    \multirow{3}{*}{Diet} & mixed & 14 & 14.4\%\\
                          & exclusive formula & 35 & 36.1\%\\
                          & exclusive breast & 48 & 49.5\%\\\hline
    \multirow{2}{*}{Delivery} & Cesarean & 98 & 30.5\%\\
                              & Vaginal & 223 & 69.5\%\\
    \hline\hline    
\end{tabular}
\caption{
    General demographics
}
\label{tab:demographics}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
    \hline\hline
    \textbf{ } & \textbf{Mean} & \textbf{Median} & \textbf{Minimum} & \textbf{Maximum} & \textbf{Total} \\\hline
    Infancy & 0.44 & 0.32 & 0.08 & 1.0 & 129 \\
    Early childhood & 3.23 & 3.03 & 1.0 & 5.99 & 198 \\
    Middle childhood & 8.61 & 8.38 & 6.01 & 11.9 & 155 \\
    Adolecence & 13.14 & 13.08 & 12.31 & 15.29 & 11 \\\hline\hline
\end{tabular}
\caption{
    Age statistics of children
}
\label{tab:agestats}
\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{../figures/05_data_summaries}
\caption{
    Graphical representations of subject and sample-associated metadata.
    (a) Ages of children when samples were collected.
    (b) Age-normalized BMI z-score of children when samples were collected.
    (c) Age-normalized cognitive score of children when samples were collected.
    (d) Subjects with longitudinal timepoints by first collection.
}
\label{fig:data}
\end{figure}
